Left-linear rules in adhesive categories

### 2024-02-14 icalp/fusione.tex
varie cose minori

\begin{restatable}[No need of useless switchings]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
\end{restatable}

@@

\begin{restatable}[Canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transposition $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.\todo{P: $\nu_{i,n} = \sigma \circ \nu_{1,i}$, giusto? forse e' piu' carino}
\end{restatable}

@@

\begin{lemma}
  \label{le:switchC}
  Let $\der{D}_i$ for $i \in \interval[0]{n+1}$ be derivation sequences
  and consider a switching sequence
  \begin{center}
    $\der{D}_0 \shift{\nu_0} \der{D}_1 \shift{\nu_1} \ldots
     \der{D}_n \shift{\nu_n} \der{D}_{n+1}$
  \end{center}
  Assume that the subsequence starting from $\der{D}_1$ consists of
  inversions and $\nu_0=\transp{i}$ while
  $\nu_n = \transp{\nu_{1,n}(i)}$.  Then there is shorter switching sequence
  consisting only of inversions:
  \begin{center}
    $\der{D}_0 \shift{\nu_2'} \der{D}_3' \shift{\nu_3'} \ldots
    \shift{\nu_{n-1}'} \der{D}_n'$
  \end{center}
\end{lemma}

@@

  Now, since switchings $\transp{i}$ twice one obtains a derivation
  which is abstraction equivalent to the starting one.  Hence we have
  \begin{center}
    $\der{D}_0 \equiv_a
    \der{D}_2' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_{n+1}$
  \end{center}
  as desired.
\end{proof}

@@

\begin{proof}
  Take a maximal subsequence
  $\der{D}_i \shift{\nu_{i+1}} \der{D}_{i+1} \shift{\nu_{i+2}} \ldots
  \shift{\nu_j}\der{D}_j$ consisting only of inversions.
  %
  If $i=0$ and $j=n$, there is nothing to show. Hence assume that
  $i>0$ and consider the sequence
  \begin{center}
    $\der{D}_{i-1} \shift{\nu_{i}} \der{D}_{i} \shift{\nu_{i+1}} \ldots
    \shift{\nu_j}\der{D}_j$
  \end{center}
  Then, by maximality, if $\nu_{i-1}=\transp{k}$ it must hold that
  $\nu_{j}=\transp{\nu_{i,j}(i)}$. Then, by Lemma~\ref{le:switchC},
  the switching sequence from $\der{D}_{i-1}$ to $\der{D}_{j}$ can be
  shortened by $2$:
  \begin{center}
    $\der{D}_{i-1} = \der{D}_{i+1}' \shift{\nu_{i+1}'} \ldots
    \shift{\nu_j'}\der{D}_j$
  \end{center}
  and thus the same applies to the original switching sequence.
\end{proof}

@@

\begin{proof}
  Let $\der{D} \shifteq \der{D}'$. Then there is there is a switching
  sequence
  \begin{center}
    $\der{D} = \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n=\der{D}'$
  \end{center}
  If it does not consist only of inversions, by Lemma~\ref{le:switchD}
  we can reduce its length by $2$. An inductive reasoning allows us to
  conclude.
\end{proof}

>>>>

\begin{restatable}[No need of useless switchings]{theorem}{thNoNeed}
  \label{th:no-need}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If $\der{D} \shifteq \der{D}'$ then there is a switching sequence consisting only of inversions
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a \der{D}'$
  \end{center}
\end{restatable}

@@

\begin{restatable}[Canonical form]{corollary}{coCanonical}
  \label{co:canonical}
  Let $\der{D}$ and $\der{D}'$ be derivation sequences. If
  $\der{D} \shifteq[\sigma]\der{D}'$ then there is a switching sequence
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a\der{D}'$
  \end{center}
  where for each $i \in \interval{n}$ the transposition $\nu_i  = \transp{k}$ is such that 
  $k = \max \{ j \mid \transp{j} \in \inv{\nu_{i,n}}\}$.\todo{P: $\nu_{i,n} = \sigma \circ \nu_{1,i}$, giusto? forse e' piu' carino}
\end{restatable}

@@

\begin{lemma}
  \label{le:switchC}
  Let $\der{D}_i$ for $i \in \interval[0]{n+1}$ be derivation sequences
  and consider a switching sequence
  \begin{center}
    $\der{D}_0 \shift{\nu_0} \der{D}_1 \shift{\nu_1} \ldots
     \der{D}_n \shift{\nu_n} \der{D}_{n+1}$
  \end{center}
  Assume that the subsequence starting from $\der{D}_1$ consists of
  inversions and $\nu_0=\transp{i}$ while
  $\nu_n = \transp{\nu_{1,n}(i)}$.  Then there is shorter switching sequence
  consisting only of inversions:
  \begin{center}
    $\der{D}_0 \equiv_a \der{D}_0' \shift{\nu_2'} \der{D}_3' \shift{\nu_3'} \ldots
    \shift{\nu_{n-1}'} \der{D}_n'$
  \end{center}
\end{lemma}

@@

  Now, switching $\transp{i}$ twice one obtains a derivation
  which is abstraction equivalent to the starting one.  Hence we have
  \begin{center}
    $\der{D}_0 \equiv_a
    \der{D}_2' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_{n+1}$
  \end{center}
  as desired.
\end{proof}

@@

\begin{proof}
  Take a maximal subsequence
  $\der{D}_i \shift{\nu_{i+1}} \der{D}_{i+1} \shift{\nu_{i+2}} \ldots
  \shift{\nu_j}\der{D}_j$ consisting only of inversions.
  %
  If $i=0$ and $j=n$, there is nothing to show. Hence assume that
  $i>0$ and consider the sequence
  \begin{center}
    $\der{D}_{i-1} \shift{\nu_{i}} \der{D}_{i} \shift{\nu_{i+1}} \ldots
    \shift{\nu_j}\der{D}_j$
  \end{center}
  Then, by maximality, if $\nu_{i-1}=\transp{k}$ it must hold that
  $\nu_{j}=\transp{\nu_{i,j}(i)}$. Then, by Lemma~\ref{le:switchC},
  the switching sequence from $\der{D}_{i-1}$ to $\der{D}_{j}$ can be
  shortened by $2$, i.e., 
  \begin{center}
    $\der{D}_{i-1} \equiv_a \der{D}_{i+1}' \shift{\nu_{i+1}'} \ldots
    \shift{\nu_j'}\der{D}_j$
  \end{center}
  and thus the same applies to the original switching sequence.
\end{proof}

@@

\begin{proof}
  Let $\der{D} \shifteq \der{D}'$. Then there is there is a switching
  sequence
  \begin{center}
    $\der{D} \equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n \equiv_a \der{D}'$
  \end{center}
  If it does not consist only of inversions, by Lemma~\ref{le:switchD}
  we can reduce its length by $2$. An inductive reasoning allows us to
  conclude.
\end{proof}
